---
title: "IS 605 Assignment 9"
author: "Oluwakemi Omotunde"
date: "April 5, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This week, we'll empirically verify Central Limit Theorem. We'll write code to run a
small simulation on some distributions and verify that the results match what we expect
from Central Limit Theorem. Please use R markdown to capture all your experiments and
code. Please submit your Rmd file with your name as the 

(1) First write a function that will produce a sample of random variable that is distributed as follows:

$$f(x) = x, 0 \leq x \leq 1$$
$$f(x) = 2 - x, 1 < x \leq 2$$

```{r prob 1}
#we will use the inverse sampling method 
samp.func1 <- function(y){
  if(y >= 0 && y <=2){
    #f(x) = x^2/2 - x + 1, x E [0,1)
    #     = 2x-x^2/2-1, c E [1, 2)
  return <- ifelse(y < .5, sqrt(2*y), 2- sqrt(2*(1-y)))
  }
}
```

```{r prob 2}
samp.func2 <- function(y){
  if(y >= 0 && y <=2){
    #f(x) = x^2/2 - x + 1, x E [0,1)
    #     = 2x-x^2/2-1, c E [1, 2)
  return <- ifelse(y < .5, sqrt(2*y), 2- sqrt(2*(1-y)))
  }
}
```

