---
title: "IS 605 Assignment 8"
author: "Oluwakemi Omotunde"
date: "April 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In the Grade Network that we looked at in the notes, what happens to the probability of
Diffculty of Course when you present the evidence that the received recommendation let-
ter was good? In addition, now present the evidence that both SAT scores were good and
the letter of recommendation was good, What is the probability of the Diffculty of Course now? You should use the gRain package in R to build your network and perform these calculations. You may need to install RBGL package from BioConductor in R to get gRain working. See http://www.bioconductor.org/packages/release/bioc/html/RBGL.html for
instructions on RBGL. Please submit your assignment as an R markdown document.

```{r set variables}
require(gRain)
#Now to set the variables going off of the chart on page 3 of our notes

level_ny  <- c("no", "yes")
level_lh <- c("low", "high")

#now to set the provided probabilities for wach category

d <- cptable(~difficulty, values = c(.3, .7), levels = level_ny)
i <- cptable(~intelligence, values = c(.8, .2), levels = level_lh)
s <- cptable(~SAT|intelligence, values = c(.9, .2, .1, .8), levels = level_lh)
g <- cptable(~grade|difficulty:intelligence, values = c(.6, .4, .8, .2, .1, .99, .1, .9), levels = level_lh)
l <- cptable(~letter|grade, values = c(.9, .01, .05, .95), levels = level_lh)
plist <- compileCPT(list(d, i, s, g, l))
#as a check I want to ake sure that my input correlates to what is provided in the reading
plist$difficulty
plist$intelligence
plist$SAT
plist$grade
plist$letter
```

Now that we have our variables as it is in the output is similar to and probablities set we can find the requested probabilities. 

```{r prob}
grain_list <-  grain(plist)
querygrain(grain_list, c(nodes = "difficulty", "intelligence", "SAT", "grade", "letter", type = "marginal"))
```
  + What happens to the probability of Diffculty of Course when you present the evidence that the received recommendation letter was good?

```{r prob diff}
prob_diff <- setEvidence(grain_list, nodes = "letter", states = "high")
querygrain(prob_diff, nodes = "difficulty", type = "marginal")
```
      + Given the evidence above, we see that the probability of a class being difficult decreases to .61 while the probability of a class not being difficult increases to .39
    
  + Now present the evidence that both SAT scores were good and the letter of recommendation was good, what is the probability of the Diffculty of Course now?
  
```{r sat diff}
prob_sat <- setEvidence(grain_list, nodes = c("SAT", "letter"), states = c("high", "high"))
querygrain(prob_sat, nodes = "difficulty", type = "marginal")
```

    + Given the above, the probability of a difficult class increased to .67 and decreased to .32